---
title: "Eelgrass table rarefied 4660 sequences per sample to generate barplot"
author: "DSB"
output: pdf_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, include = TRUE, fig.show=TRUE) #global options
```

Package versions for phyloseq, ggplot2, plyr, data.table, tidyverse, knitr
```{r, echo=FALSE, message=FALSE}
library(phyloseq); packageVersion("phyloseq") 
library(ggplot2); packageVersion("ggplot2")
library(plyr); packageVersion("plyr")
library(data.table); packageVersion("data.table")
library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion(("knitr"))
gsub("[[:punct:][:space:]]", "-", Sys.time())
```

```{r read in data}
Eelgrass4660PhyloseqObj <- readRDS(file = "../../phyloseq_object/Eelgrass4660PhyloseqObj.rds")

#order factor levels
sample_data(Eelgrass4660PhyloseqObj)$RegionName <-factor(sample_data(Eelgrass4660PhyloseqObj)$RegionName, levels=c("Alaska", "British Columbia", "Washington", "Oregon", "Bodega", "San Diego"))

#order factor levels
sample_data(Eelgrass4660PhyloseqObj)$Region <- factor(sample_data(Eelgrass4660PhyloseqObj)$Region, levels=c("AK", "BC", "WA", "OR", "BB", "SD"))

```

Summary of phyloseq object
```{r summarize}
Eelgrass4660PhyloseqObj #393 samples
```

``````{r barplot order, warning=FALSE}
#merge ESVs that have the same classification at taxonomic rank Order
Eelgrass4660PhyloseqObjGlomOrder = tax_glom(Eelgrass4660PhyloseqObj, taxrank = "Order", NArm = FALSE)
# print phyloseq object summary
Eelgrass4660PhyloseqObjGlomOrder # 200 taxa 393 samples
# check sample sums equal 4660 seqs/sample
sample_sums(Eelgrass4660PhyloseqObjGlomOrder)

# psmelt to melt wide to long format of taxonomy table and merge with dataframe of metadata
Eelgrass4660PhyloseqObjGlomOrderdf = psmelt(Eelgrass4660PhyloseqObjGlomOrder)

# vector of tableau20 hexidecimal codes for colors
col2 =  c("#4E79A7", "#A0CBE8", "#F28E2B", "#FFBE7D", "#59A14F", "#8CD17D", "#B6992D", "#F1CE63", "#499894", "#86BCB6", "#E15759", "#FF9D9A", "#79706E", "#BAB0AC", "#D37295", "#FABFD2", "#B07AA1", "#D4A6C8", "#9D7660", "#D7B5A6")

# print unique factor levels
unique(Eelgrass4660PhyloseqObjGlomOrderdf$TissueType)
# rename levels of factor for plot
Eelgrass4660PhyloseqObjGlomOrderdf$TissueType <- gsub("green tissue lesioned blade", "GTLB", Eelgrass4660PhyloseqObjGlomOrderdf$TissueType)
Eelgrass4660PhyloseqObjGlomOrderdf$TissueType <- gsub("green tissue non-lesioned blade", "non-LB", Eelgrass4660PhyloseqObjGlomOrderdf$TissueType)
Eelgrass4660PhyloseqObjGlomOrderdf$TissueType <- gsub("lesion tissue", "Lesion", Eelgrass4660PhyloseqObjGlomOrderdf$TissueType)

# subset dataframe by columns of interest
Eelgrass4660PhyloseqObjGlomOrderdfSubset <- subset(Eelgrass4660PhyloseqObjGlomOrderdf, select = c(Abundance, Order, Region, TissueType))

# create three separate dataframes for each tissue type
# Lesion tissue dataframe
Eelgrass4660PhyloseqObjGlomOrderdfSubsetLesion <- Eelgrass4660PhyloseqObjGlomOrderdfSubset %>% filter(TissueType == "Lesion")

# check that you correctly filtered dataset
unique(Eelgrass4660PhyloseqObjGlomOrderdfSubsetLesion$TissueType)
nrow(Eelgrass4660PhyloseqObjGlomOrderdfSubsetLesion)

# GTLB dataframe
Eelgrass4660PhyloseqObjGlomOrderdfSubsetGTLB <- Eelgrass4660PhyloseqObjGlomOrderdfSubset %>% filter(TissueType == "GTLB")

# check that you correctly filtered dataset
unique(Eelgrass4660PhyloseqObjGlomOrderdfSubsetGTLB$TissueType)
nrow(Eelgrass4660PhyloseqObjGlomOrderdfSubsetGTLB)

# non-LB dataframe
Eelgrass4660PhyloseqObjGlomOrderdfSubsetnonLB <- Eelgrass4660PhyloseqObjGlomOrderdfSubset %>% filter(TissueType == "non-LB")

# check that you correctly filtered dataset
unique(Eelgrass4660PhyloseqObjGlomOrderdfSubsetnonLB$TissueType)
nrow(Eelgrass4660PhyloseqObjGlomOrderdfSubsetnonLB)

# calculate relative abundances of each Order per region on each tissue type dataframe
# loop through list of unique region names to subset the df; group_by Order and sum abundances of each Order, calculate sum of counts for each region (N), calculate relative abundances of each Order, check that sum of relative abundances per region equal 100%, append dataframe to a list, bind dataframes within list with bind_rows function

# obtain unique list of RegionNames
region_unique <- unique(Eelgrass4660PhyloseqObjGlomOrderdfSubset$Region)

# create list for lesion data
region_table_lesion <- list()

# for loop for lesion blade dataframe
for(i in seq_along(region_unique)) {
  X=subset(Eelgrass4660PhyloseqObjGlomOrderdfSubsetLesion, Region == region_unique[i])
  tempLesion = X %>% group_by(Order) %>% summarise_each(funs(sum (Abundance)))
  tempLesion$Region = region_unique[i]
  tempLesion$TissueType = c("Lesion")
  tempLesion$N =  sum(tempLesion$Abundance) 
  tempLesion$RelAbund <- ( (tempLesion$Abundance / tempLesion$N) * 100 )
  tempLesion$SumRelAbund <- sum(tempLesion$RelAbund) 
  region_table_lesion[[i]] <- tempLesion
  
  print(tempLesion)
}

# bind items in region_table list with rbind/bind_rows
region_table_lesion <- bind_rows(region_table_lesion)

# create list for GTLB data
region_table_GTLB <- list()

# for loop for GTLB dataframe
for(i in seq_along(region_unique)) {
  X=subset(Eelgrass4660PhyloseqObjGlomOrderdfSubsetGTLB, Region == region_unique[i])
  tempGTLB = X %>% group_by(Order) %>% summarise_each(funs(sum (Abundance)))
  tempGTLB$Region = region_unique[i]
  tempGTLB$TissueType = c("GTLB")
  tempGTLB$N =  sum(tempGTLB$Abundance) 
  tempGTLB$RelAbund <- ( (tempGTLB$Abundance / tempGTLB$N) * 100 )
  tempGTLB$SumRelAbund <- sum(tempGTLB$RelAbund) 
  region_table_GTLB[[i]] <- tempGTLB
  
  print(tempGTLB)
}

# bind items in region_table list with rbind/bind_rows
region_table_GTLB <- bind_rows(region_table_GTLB)

# create list for non-LB data
region_table_nonLB <- list()

# for loop for non-LB dataframe
for(i in seq_along(region_unique)) {
  X=subset(Eelgrass4660PhyloseqObjGlomOrderdfSubsetnonLB, Region == region_unique[i])
  tempnonLB = X %>% group_by(Order) %>% summarise_each(funs(sum (Abundance)))
  tempnonLB$Region = region_unique[i]
  tempnonLB$TissueType = c("non-LB")
  tempnonLB$N =  sum(tempnonLB$Abundance) 
  tempnonLB$RelAbund <- ( (tempnonLB$Abundance / tempnonLB$N) * 100 )
  tempnonLB$SumRelAbund <- sum(tempnonLB$RelAbund) 
  region_table_nonLB[[i]] <- tempnonLB
  
  print(tempnonLB)
}

# bind items in region_table list with rbind/bind_rows
region_table_nonLB <- bind_rows(region_table_nonLB)

# bind each tissue type dataframe
region_table_all <- bind_rows(region_table_lesion,region_table_GTLB,region_table_nonLB)

# create column of top twenty most abundant orders across all regions

#subset to Order and relative abundance column 
region_table_subset <-  subset(region_table_all, select = c(Order, RelAbund))

# group by Order and calculate mean relative abundances across all regions
RegionalMeanRelAbundOrder <- region_table_subset %>%
  group_by(Order) %>%
  summarise_each(funs(mean (RelAbund)))

# sort by decreasing Order mean relative abundance 
RegionalMeanRelAbundOrder <- RegionalMeanRelAbundOrder[order(RegionalMeanRelAbundOrder$RelAbund, decreasing = TRUE), ]
# subset by the first 20 rows (~ top twenty most abundant Orders)
top_twenty <- RegionalMeanRelAbundOrder[1:21, ]
# remove Order NA
top_twenty <- top_twenty %>% filter(Order != "NA")

# obtain list of top twenty order names
 top_twenty_names <- top_twenty$Order

# create new column of top twenty most abundant Orders in region_table dataframe
region_table_all$Order20 <- NA

# fill Order20 column with Order column names of 20 most abundant Orders using ifelse statement
region_table_all$Order20 <- ifelse(region_table_all$Order %in% top_twenty_names, region_table_all$Order, NA)

#order factor levels
region_table_all$TissueType <- factor(region_table_all$TissueType, levels=c("non-LB", "GTLB", "Lesion"))

# bar plot of top 20 most abundant Orders, with full metadata and facet grid 
ggplot(region_table_all, aes(x = Region, y = RelAbund, fill = Order20)) + 
  geom_bar(stat="identity") +
  scale_fill_manual(values = col2) +
  theme(legend.title = element_text(size=16), legend.text = element_text(size=14), axis.text=element_text(size=14), axis.title = element_text(size = 18)) +
  theme(strip.text.x = element_text(size = 14)) +
  facet_grid(~TissueType) +
  ylab("Relative abundance %") +
  xlab("Region") +
  guides(fill=guide_legend("Order"))

ggsave(file = "../../plots/Eelgrass_4660_barplot_top20_Order.png", 
       dpi = 1000, device = "png", width = 11, height = 8.5, units = "in")
```

