---
title: "Microbial analyses on water samples rarefied to 4660 sequences"
author: "DSB"
output: github_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(eval = TRUE, echo = FALSE, include = TRUE, fig.show=TRUE) #global options
```

Package versions for phyloseq, vegan, ggplot2, plyr, data.table, tidyverse, nlme, ggpubr, ggsci, scales

```{r, echo=FALSE, message=FALSE}
library(phyloseq); packageVersion("phyloseq")
library(vegan); packageVersion("vegan")
library(ggplot2); packageVersion("ggplot2")
library(plyr); packageVersion("plyr")
library(data.table); packageVersion("data.table")
library(tidyverse); packageVersion("tidyverse")
library(nlme); packageVersion("nlme")
library(ggpubr); packageVersion("ggpubr")
library(ggsci); packageVersion("ggsci")
library(scales); packageVersion("scales")

gsub("[[:punct:][:space:]]", "-", Sys.time())
```

```{r read in data, results=FALSE}
Water4660PhyloseqObj <- readRDS(file = "../../phyloseq_object/Water4660PhyloseqObj.rds")

# order factor levels
sample_data(Water4660PhyloseqObj)$RegionName <- factor(sample_data(Water4660PhyloseqObj)$RegionName, levels=c("Alaska", "British Columbia", "Washington", "Oregon", "Bodega", "San Diego"))

# order factor levels
sample_data(Water4660PhyloseqObj)$Region <- factor(sample_data(Water4660PhyloseqObj)$Region, levels=c("AK", "BC", "WA", "OR", "BB", "SD"))

# print phyloseq summary
Water4660PhyloseqObj #92 samples includes extra samples that are not part of the wasting disease project

#remove Drake's Estero samples
Water4660PhyloseqObj <- prune_samples(sample_names(Water4660PhyloseqObj) != "Sample36W", Water4660PhyloseqObj)
Water4660PhyloseqObj <- prune_samples(sample_names(Water4660PhyloseqObj) != "Sample61W", Water4660PhyloseqObj)
Water4660PhyloseqObj <- prune_samples(sample_names(Water4660PhyloseqObj) != "Sample84W", Water4660PhyloseqObj)
#remove Sacramento Landing samples
Water4660PhyloseqObj <- prune_samples(sample_names(Water4660PhyloseqObj) != "Sample52W", Water4660PhyloseqObj)
Water4660PhyloseqObj <- prune_samples(sample_names(Water4660PhyloseqObj) != "Sample57W", Water4660PhyloseqObj)
Water4660PhyloseqObj <- prune_samples(sample_names(Water4660PhyloseqObj) != "Sample73W", Water4660PhyloseqObj)

```

Summary of phyloseq object, number of samples, number of taxa, sample variables, taxa levels from water samples after removing Drake's Estero and Sacramento Landing samples (sites that are not part of the wastin disease study). There are 86 samples and 15,418 exact sequence variants. 

```{r summarize}
Water4660PhyloseqObj

```

Betadiversity analyses and PCoA plots on Bray Curtis dissimilarity after removing samples from sites that lack SST records (reduces sample size to 71). 

```{r betadiv, warning=FALSE}

# extract metadata from phyloseq object into a df
Metadata <- data.frame(sample_data(Water4660PhyloseqObj))

# remove samples with missing metadata of interest (e.g. satellite temps via MUR & GHRSST)
Water4660PhyloseqObjSatelliteTempNARemoved <- Water4660PhyloseqObj %>%
  subset_samples(!is.na (SST))

# print phyloseq summary
Water4660PhyloseqObjSatelliteTempNARemoved # 71 samples

# extract metadata phyloseq obj following removal of NA for SST data
Metadata.2 <- data.frame(sample_data(Water4660PhyloseqObjSatelliteTempNARemoved))

# Bray Curtis dissimilarity
Water4660PhyloseqObjSatelliteTempNARemovedDistanceBrayCurtis = distance(Water4660PhyloseqObjSatelliteTempNARemoved, method = "bray")

#ordination PCoA
OrdinationWater4660DistanceBrayCurtisTempNARemoved = ordinate(Water4660PhyloseqObjSatelliteTempNARemoved, method = "PCoA", 
distance = Water4660PhyloseqObjSatelliteTempNARemovedDistanceBrayCurtis)

# adonis on Bray Curtis dissimilarity with meadow location nested within region
BrayCurtisWater4660Adonis <- adonis2(Water4660PhyloseqObjSatelliteTempNARemovedDistanceBrayCurtis ~ PrevalenceMean + SeverityMean + BladeAreaMean + DensityShootsMean + SheathLengthMean + LongestBladeLengthMean + SST + thirty_day_max + thirty_day_min + Region/LocationName, by = "terms", data = Metadata)

BrayCurtisWater4660Adonis
write.csv(BrayCurtisWater4660Adonis, "../../out_files/BrayCurtisWater4660Adonis2LocationNameNestedRegion.csv")

# obtain color palette for plot
mypalNEJM <- pal_nejm("default")(8)
# plot colors 
show_col(mypalNEJM)
# assign hexidecimal color codes to variable
mypalNEJM =  c("#BC3C29FF", "#0072B5FF", "#E18727FF", "#20854EFF", "#7876B1FF", "#6F99ADFF") 

#plot by SST
Water_4660_Bray_Curtis_PCoA_Region_SST_NA_removed <- plot_ordination(Water4660PhyloseqObjSatelliteTempNARemoved,OrdinationWater4660DistanceBrayCurtisTempNARemoved, color = "RegionName") + 
  geom_point(mapping = aes(size = SST)) +
  scale_colour_manual(values = mypalNEJM) +
  theme(legend.title = element_text(size=14), legend.text = element_text(size=14), axis.text=element_text(size=14), axis.title = element_text(size = 14) ) +
  labs(color = "Region", size = "SST ºC") +
  guides(color = guide_legend(override.aes = list(size=6)))

Water_4660_Bray_Curtis_PCoA_Region_SST_NA_removed

ggsave(Water_4660_Bray_Curtis_PCoA_Region_SST_NA_removed, file = "../../plots/Water_4660_Bray_Curtis_PCoA_Region_SST_NA_removed.png", 
       dpi = 1000, device = "png", width = 11, height = 8.5, units = "in")

# use ggarange to put PCoA of eelgrass green 4660 samples and water 4660 samples into one figure
# Green_only_4660_Bray_Curtis_PCoA_Region_SST_thirty_day_max comes from phyloseq_green4660.Rmd code, load object into environment

# ggarange eelgrass and water PCoAs
PCoA_Figure <- ggarrange(Green_only_4660_Bray_Curtis_PCoA_Region_SST_thirty_day_max, Water_4660_Bray_Curtis_PCoA_Region_SST_NA_removed,
                 labels = c("A", "B"), ncol = 1, nrow = 2, align = "hv", heights = c(1.4,1))

PCoA_Figure

ggsave(PCoA_Figure, file = "../../plots/Wrap_PCoA_Green_only_4660_Water_4660_Bray_Curtis_SST.png", 
       dpi = 1000, device = "png", width = 11, height = 8.5, units = "in")

``` 

Test for dispersion differences in seawater microbiome composition with region.

```{r betadisp & visualize, warning=FALSE, message=FALSE}
#calculate multivariate dispersions
BetadispWater4660Region <- (betadisper(Water4660PhyloseqObjSatelliteTempNARemovedDistanceBrayCurtis, Metadata.2$RegionName))
# permtest of dispersions
permutest(BetadispWater4660Region, pairwise = TRUE)
BetadispWater4660Region
capture.output(BetadispWater4660Region, file = "../../out_files/BetadispWater4660Region.txt")
capture.output(permutest(BetadispWater4660Region, pairwise = TRUE), file = "../../out_files/BetadispWater4660RegionPermTest.txt")

```

Alphadiversity analyses on seawater microbial communties after removing samples from sites that lack SST records (reduces sample size to 71)

```{r alphadiv}
#calculate alpha diversity metrics, richness & Shannon diversity
AlphaWater4660 <- estimate_richness(Water4660PhyloseqObj, measures = c("Observed", "Shannon"))

# plot histograms
histogram(AlphaWater4660$Observed)
histogram(AlphaWater4660$Shannon)

#make SampleID as factor
AlphaWater4660$SampleID <- rownames(AlphaWater4660) %>% as.factor()

#extract metadata from phyloseq object into a df, make SampleID as factor
metadata <- data.frame(sample_data(Water4660PhyloseqObj))
metadata$SampleID <- rownames(metadata) %>% as.factor()

# left join metadata with Alpha diversity by SampleID
MetadataAlphaDivWater <- metadata %>% 
  unclass() %>% 
  data.frame() %>% 
  left_join(AlphaWater4660, by = "SampleID")

# remove samples without SST data
MetadataAlphaDivWaterSSTNARemoved <- MetadataAlphaDivWater %>% 
  filter(SST != "NA")

# mixed effect model on richness (observed ESVs), meadow location as a random effect
m1 <-  lme(Observed ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "REML", data = MetadataAlphaDivWaterSSTNARemoved)

# summaries and anova on model
m1
anova(m1)
summary(m1)

# test for heterogeneity of variance  
E2 <- resid(m1, type = "normalized")
F2 <- fitted(m1)
op <- par(mfrow = c(2,2), mar = c(4, 4, 3, 2))
MyYlab <- "Residuals"
plot(x = F2, y = E2, xlab = "Fitted values", ylab = MyYlab) #pattern of variance 

# plot residuals by explanatory factors
boxplot(E2 ~ SeverityMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "Lesion severity", ylab = MyYlab) 
boxplot(E2 ~ PrevalenceMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "Lesion prevalence", ylab = MyYlab)
boxplot(E2 ~ DensityShootsMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "Shoot density", ylab = MyYlab)
boxplot(E2 ~ SheathLengthMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "Sheath length Mean", ylab = MyYlab)
boxplot(E2 ~ RegionName, data = MetadataAlphaDivWaterSSTNARemoved, main = "Region", ylab = MyYlab) 
boxplot(E2 ~ SST, data = MetadataAlphaDivWaterSSTNARemoved, main = "SST", ylab = MyYlab) 

# model the variance 
# VarExp can take the value of 0
vf <- varComb(varExp(form=~SeverityMean), varExp(form=~SST))

# use varComb varExp for SeverityMean and SST in mixed effects model
# check residuals versus fitted values and explanatory variables above
m1 <-  lme(Observed ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "REML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)

# model summaries and anova
m1
anova(m1)
summary(m1)

#sequential drop of factors and compare model AIC scores with anova on ML
m1.1 <-  lme(Observed ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "ML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)
anova(m1)

# drop SST variable and compare to m1.1 full model
m2 <-  lme(Observed ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName,
random = ~1 | LocationName, method = "ML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)

# compare models
anova(m1.1,m2) # lower AIC but not significantly different
anova(m2)

# keep m1 full model with variance structure modelled with REML method
m1 <-  lme(Observed ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "REML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)

# model summaries and anova
m1
anova(m1)
summary(m1)

# correlation test of observed sequence variant richness with severity mean and sheath lenth mean
with(MetadataAlphaDivWaterSSTNARemoved, cor.test(SeverityMean, Observed)) # - 0.32 p = 0.006
with(MetadataAlphaDivWaterSSTNARemoved, cor.test(SheathLengthMean, Observed)) # 0.72 p < 0.001

# save anova of m1 model as data frame 
m1observed_anova <- anova(m1)
m1observed_anova
Alpha_div_observed_water_4660_lme_LocationName_random_effect <- as.data.frame(m1observed_anova)

write_csv(Alpha_div_observed_water_4660_lme_LocationName_random_effect, "../../out_files/Alpha_div_observed_water_4660_lme_LocationName_random_effect.csv")

# shannon mixed effects model
m1shannon <-  lme(Shannon ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "REML", data = MetadataAlphaDivWaterSSTNARemoved)

# model summary and anova
m1shannon
anova(m1shannon)
summary(m1shannon)

#test for heterogeneity of variance  
E2 <- resid(m1shannon, type = "normalized")
F2 <- fitted(m1shannon)
op <- par(mfrow = c(2,2), mar = c(4, 4, 3, 2))
MyYlab <- "Residuals"
plot(x = F2, y = E2, xlab = "Fitted values", ylab = MyYlab) #pattern of variance 

boxplot(E2 ~ SeverityMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "SeverityMean", ylab = MyYlab) 
boxplot(E2 ~ PrevalenceMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "PrevalenceMean", ylab = MyYlab) 
boxplot(E2 ~ DensityShootsMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "Shoot density", ylab = MyYlab)
boxplot(E2 ~ SheathLengthMean, data = MetadataAlphaDivWaterSSTNARemoved, main = "Sheath Length Mean", ylab = MyYlab)
boxplot(E2 ~ RegionName, data = MetadataAlphaDivWaterSSTNARemoved, main = "Region", ylab = MyYlab)
boxplot(E2 ~ SST, data = MetadataAlphaDivWaterSSTNARemoved, main = "SST", ylab = MyYlab) 

# use varComb varExp for SeverityMean and SST in mixed effects model
# check residuals versus fitted values and explanatory variables above
vf <- varComb(varExp(form=~SeverityMean), varExp(form=~SST)) 

#shannon mixed effects model with variance structure
m1shannon <-  lme(Shannon ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "REML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)

# model summary and anova
m1shannon
anova(m1shannon)
summary(m1shannon)

#sequential drop of factors to compare AIC and anova of ML method; prevalance has lowest F value
m1shannon <-   lme(Shannon ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "ML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)
anova(m1shannon)

# drop factor with lowest F value, prevalence mean
m1shannon.1 <-   lme(Shannon ~ SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "ML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)
# compare models
anova(m1shannon, m1shannon.1) # NSD

# best model is with all factors and weighted variance structure exponential for SeverityMean and SST
m1shannon <-  lme(Shannon ~ PrevalenceMean + SeverityMean + DensityShootsMean + SheathLengthMean + RegionName + SST,
random = ~1 | LocationName, method = "REML", data = MetadataAlphaDivWaterSSTNARemoved, weights = vf)

# model summary and anova
m1shannon
anova(m1shannon)
summary(m1shannon)

# correlation test of sequence variant shannon diversity with severity mean and sheath lenth mean
with(MetadataAlphaDivWaterSSTNARemoved, cor.test(SeverityMean, Shannon))  # -0.24 p = 0.042
with(MetadataAlphaDivWaterSSTNARemoved, cor.test(SheathLengthMean, Shannon)) # 0.67 p < 0.001

# as dataframe
m1shannon_anova <- anova(m1shannon)
Alpha_div_shannon_diversity_water_4660_lme_LocationName_random_effect <- as.data.frame(m1shannon_anova)

write_csv(m1shannon_anova, file = "../../out_files/Alpha_div_shannon_water_4660_LME_LocationName_random_effect.csv")

# assign hexidecimal color codes to variable to match beta diversity colors
mypalNEJM =  c("#BC3C29FF", "#0072B5FF", "#E18727FF", "#20854EFF", "#7876B1FF", "#6F99ADFF") 

# plot by richness SeverityMean 
B <- 
  ggplot(MetadataAlphaDivWaterSSTNARemoved, aes(x = SeverityMean, y = Observed, 
                                                          color = RegionName)) + 
  geom_jitter(size = 5, width = 0.1) +
  ylab("Richness") +
  xlab("Lesion severity per transect") + 
  theme(text = element_text(size=16)) +
  scale_colour_manual(values = mypalNEJM, name="Region") 

# plot Shannon diversity by SeverityMean
D <- 
  ggplot(MetadataAlphaDivWaterSSTNARemoved, aes(x = SeverityMean, y = Shannon, 
                                                          color = RegionName)) + 
  geom_jitter(size = 5, width = 0.1) +
  ylab("Shannon diversity") +
  xlab("Lesion severity per transect") + 
  ylim(4,6) +
  theme(text = element_text(size=16)) +
  scale_colour_manual(values = mypalNEJM, name="Region") 

# plot richness by SheathLengthMean 
A <- ggplot(MetadataAlphaDivWaterSSTNARemoved, aes(x = SheathLengthMean, y = Observed, 
                                                          color=RegionName)) + 
  geom_jitter(size = 5, width = 0.1) +
  ylab("Richness") +
  xlab("Sheath length per transect") + 
  theme(text = element_text(size=16)) +
  scale_colour_manual(values = mypalNEJM, name="Region") 

# plot Shannon diversity by SheathLengthMean 
C <- ggplot(MetadataAlphaDivWaterSSTNARemoved, aes(x = SheathLengthMean, y = Shannon, 
                                                          color = RegionName)) + 
  geom_jitter(size = 5, width = 0.1) +
  ylab("Shannon diversity") +
  xlab("Sheath length per transect") + 
  ylim(4,6) +
  theme(text = element_text(size=16)) +
  scale_colour_manual(values = mypalNEJM, name="Region") 

# Extract the legend; returns a gtable to use in ggarange below
leg <- get_legend(C)

# use ggarange to make figure of four alpha diversity plots above (A-D)
Fig <- ggarrange(A, B, C, D, 
    labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2, align = "hv", heights = c(1,1), legend.grob = leg, legend = "right")

Fig
ggsave(file = "../../plots/Water_4660_alpha_diversity_SheathLengthMean_SeverityMean.png", dpi = 1000, device = "png", width = 11, height = 8.5, units = "in")

```

Bar plots of the top twenty most abundant orders of bacteria associated with water.

``` {r barplot, warning=FALSE}

#merge ESVs that have the same classification at taxonomic rank order
Water4660PhyloseqObjGlomOrder = tax_glom(Water4660PhyloseqObj, taxrank = "Order", NArm = FALSE)
# summary of phyloseq object
Water4660PhyloseqObjGlomOrder

# color for plots
col3 = c("#7f3b08",
"#b35806",
"#e08214",
"#fdb863",
"#fee0b6",
"#e6f5d0",
"#b8e186",
"#7fbc41",
"#4d9221",
"#276419",
"#d1e5f0",
"#92c5de",
"#4393c3",
"#2166ac",
"#053061",
"#d8daeb", #purple
"#b2abd2", #purple
"#8073ac", #purple
"#542788", #purple
"#2d004b") #purple

# psmelt to melt wide to long format taxonomy table and combine with metadata
Water4660PhyloseqObjGlomOrderdf = psmelt(Water4660PhyloseqObjGlomOrder)

# subset dataframe by columns of interest
Water4660PhyloseqObjGlomOrderdf <- subset(Water4660PhyloseqObjGlomOrderdf, select = c(Abundance, Order, Region))

# calculate relative abundances of each Order per region
# loop through list of unique region names to subset the df; group_by Order and sum abundances of each Order, calculate sum of counts for each region (N), calculate relative abundances of each Order, check that sum of relative abundances per region equal 100% 

# obtain unique list of RegionNames
region_unique <- unique(Water4660PhyloseqObjGlomOrderdf$Region)

# create list
region_table <- list()

for(i in seq_along(region_unique)) {
  X=subset(Water4660PhyloseqObjGlomOrderdf, Region == region_unique[i])
  temp = X %>% group_by(Order) %>% summarise_each(funs(sum (Abundance)))
  temp$Region = region_unique[i]
  temp$N =  sum(temp$Abundance) 
  temp$RelAbund <- ( (temp$Abundance / temp$N) * 100 )
  temp$SumRelAbund <- sum(temp$RelAbund) 
  region_table[[i]] <- temp
  
  print(temp)
}

# bind items in region_table list with rbind/bind_rows
region_table <- bind_rows(region_table)

# create column of top twenty most abundant orders across all regions

#subset to Order and relative abundance column 
region_table_subset <-  subset(region_table, select = c(Order, RelAbund))

# group by Order and calculate mean relative abundances across all regions
RegionalMeanRelAbundOrder <- region_table_subset %>%
  group_by(Order) %>%
  summarise_each(funs(mean (RelAbund)))

# sort by decreasing Order mean relative abundance 
RegionalMeanRelAbundOrder <- RegionalMeanRelAbundOrder[order(RegionalMeanRelAbundOrder$RelAbund, decreasing = TRUE), ]
# subset by the first 20 rows (~ top twenty most abundant Orders)
top_twenty <- RegionalMeanRelAbundOrder[1:20, ]

# obtain list of top twenty order names
 top_twenty_names <- top_twenty$Order

# create new column of top twenty most abundant Orders in region_table dataframe
region_table$Order20 <- NA

# fill Order20 column with Order column names using ifelse statement
region_table$Order20 <- ifelse(region_table$Order %in% top_twenty_names, region_table$Order, NA)

# bar plot of 20 most abundant Orders
ggplot(region_table, aes(x = Region, y = RelAbund, fill = Order20)) + 
  geom_bar(stat="identity") +
  scale_fill_manual(values = col3) +
  theme(legend.title = element_text(size=14), legend.text = element_text(size=14), axis.text=element_text(size=16), axis.title = element_text(size = 16)) +
  ylab("Relative abundance %") +
  xlab("Region") +
  guides(fill=guide_legend("Order"))

 ggsave(file = "../../plots/Water_only_4660_barplot_top20_Orders.png", 
        dpi = 1000, device = "png", width = 11, height = 8.5, units = "in")

```

